The importance of fidelity in the scheduling of Redox Flow Batteries.

### 2018-11-06 Paper/main.tex
Update main.tex

\usepackage{graphicx}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
%\setlength{\parindent}{0in}
\usepackage{times}
\usepackage[british]{babel}
\usepackage{algorithm}
\usepackage{algorithmic} \renewcommand{\algorithmicrequire}{\textbf{Given:}} \renewcommand{\algorithmicensure}{\textbf{Returns:}}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\renewcommand \thesection{\arabic{section}}
\usepackage[caption=false]{subfig}
\usepackage[compact]{titlesec}
\usepackage{soul}
\usepackage{textcomp}
\usepackage[section]{placeins}
\usepackage{url}
\usepackage{xspace}
\usepackage{lineno}
\usepackage{cleveref}

%Added by Diarmid
\usepackage{siunitx}
\usepackage{hyperref}
    \hypersetup{colorlinks=true,linkcolor=green}

@@

%% main text
\section{Introduction}
\label{sec:Intro}
The  growing deployment of intermittent renewable power sources \sol{you need numbers here to underline how important this is, i.e. what percentage of energy, around the world, is now renewable} such as wind and solar in industrialised nations in order to substitute fossil fuel power is a key driver of research into the development and applications of battery energy storage systems \sol{why? what does it do? the argument needs to highlight the properties that batteries have and why we think are important} (BESS). \sol{you need to pack in the references here, you mention research so we need references to know what they looked at...}

At present in the UK, the fluctuating output from these power sources is topped up to match demand primarily by varying the output of gas fired power plant \cite{DraxElectricInsights}. In order to meet the strict CO\textsubscript{2} emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030 \cite{FES2018}, it is probable that further multi-hour energy storage will be required to displace gas plant from the role of balancing.  Within the UK electrical market, the phasing out of gas power plants, likely via increments to the carbon price \sol{reference?}, to be replaced by intermittent generation would lead to high electrical prices where demand exceeds supply, and low (or negative) pricing where supply exceeds demand.

At present, revenue from arbitrage on the wholesale electrical price is not in itself sufficient to justify BESS investment \cite{Johnston2015}. and the deployment of grid scale BESS is dominated by projects that have won contracts for short term stability services tendered by National Grid. For example, \SI{500}{\mega\watt} of BESS capacity won contracts for the T-4 2020/21 capacity market (CM) auction, and \SI{200}{\mega\watt} of BESS capacity won enhanced frequency response (EFR) contracts. The majority of successful BESS applicants for the EFR auction also have CM contracts. \sol{This is a bit parochial, the same is regarding BESS elsewhere (i.e. the big Tesla battery). We need to widen the story out to make the paper relevant outside the UK}

However, due to concerns that the average storage duration (i.e. energy to power ratio) of the BESS contracted for CM is shorter than the \SI{2}{\hour} that a stress event may last for, National Grid recently applied derating factors according to the energy to power ratio. Hence, a system rated at \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} would only receive capacity payments for \SI{0.68}{\mega\watt} in the 2018/19 T-1 auction, and \SI{0.65}{\mega\watt} in the 2021/22 T-4 auction. The derating factor is relaxed to 0.96 for systems rated for 4+ hours of discharge \cite{NGDeratingFactors}.

This development disfavours the power-centric Li-ion systems that previously won the majority of EFR contracts (Cite please) and favours redox flow battery (RFB) systems, where the incremental cost of energy capacity is low, and systems are typically specified with a 3-6 \si{\hour} energy to power ratio.

In order to quantify the economic merits of RFB systems in comparison to other BESS, it is necessary to perform techno-economic analyses (TEA) for realistic applications. TEA should involve optimisation of the operation of a BESS to maximise revenue (or minimise costs). This should be done at various energy and power ratings (as these are decoupled in a non-hybrid RFB) to find the global optimum, as described by \cite{Oudalov2007}. The optimisation of operation requires determining the schedule, that is, the power output of the battery in each sub-period that maximises benefit without violating energy conservation.

Schedule optimisation is often performed by posing a linear programming (LP) problem \cite{}. In this method, the objective function and the constraints on the system are expressed as linear algebraic terms. On the resulting geometric space, the optimum is guaranteed to lie on a vertex, hence a costly search of the whole surface is not required.

@@

\section{Model formulation}
\label{sec:Model_Formulation}
Pure price arbitrage is chosen as the application for schedule optimisation due to its simplicity, allowing a focus on the treatment of the RFB within the problem. In the following section, an LP formulation for schedule optimisation for maximum arbitrage revenue is first posed in current-density terms to serve as a reference. Then the three augmentations to the model described in \cref{sec:Intro} are introduced. Finally, the RFB specification method is outlined, including the assumption regarding the rated efficiency.
\section{LP schedule Optimisation in Current-Density Terms}
\label{Model_Formulation_LP_Reference_Current_Density_Terms}
In the literature on RFB schedule optimisation, scheduling problems are posed using either power (\cite{Chen2012,Hu2010,Nguyen2015}) or energy (\cite{Johnston2015,Vaca2017}) as the time-indexed variables. This is not an important difference, as the former only requires that the timestep $\tau$ be included as a coefficient. Here, the power convention is adopted, as it allows the timestep to be varied independently. The objective function for pure arbitrage revenue described by \cite{Hu2010} is adopted here, but adapted so that separate charge and discharge power variables are explicitly defined. Specifying a single power variable with negative values representing charge and positive discharge is not practical for LP optimisation, as the solver will drive the variable to extremes which is likely to be sub-optimal.

@@

\subsection{NLP treatment of voltaic losses}
\label{Model_Formulation_NLP_treatment_voltaic_losses}
A major simplification in the LP problem in terms of power is the assumption of a constant energy efficiency parameter. This assumption also applies to the LP problem in current-density terms introduced in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, although separate voltaic and coulombic efficiencies parameters are accounted for for. In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density. The faradaic over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density (CITE article I used in CR). \cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation. This results in a non-linear (NLP) objective function for revenue in period $Y$ given by:

@@

where $V_a$ and $ASR$ are scalar parameters representing faradaic over-potential (\si{\volt}) and area specific resistance (\si{\ohm\square\meter}). The placement of $V_a$ and $ASR$ reflects an assumption of symmetry in voltaic losses between charge and discharge (discussed further in \cref{analysis_of_assumptions}). The objective function is subject to the same constraints as described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}.

Although non-linear, \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} the objective function is convex (as it requires maximising a negative quadratic), hence a gradient based algorithmic solver is guaranteed to find the optimum from any starting point.

@@

Both the LP and NLP solutions involve charging the VRFB when the electrical price is low (e.g. between 05:00 and 07:00) and discharging it when the price is high (e.g. 10:00 to 13:00). The LP optimal profile involves discharging or charging the system at high current-density in the four hours with the most extreme prices. The NLP optimal schedule also targets the most extreme periods, but spreads the charging/discharging out. This is the consequence of factoring variable ohmic losses in the objective function, where high current-density is penalised by higher losses. Under the LP formulation, the VRFB is predicted to generate \pounds 0.068, whereas the NLP formulation predicts a revenue of \pounds 0.063. In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds 0.051. Across 2017, LP scheduling would have resulted in \pounds 22.61, against a prediction of \pounds 33.16, and NLP scheduling would have yielded \pounds 27.99.

@@

    \item Coulombic efficiency is assumed to be constant with current-density, i.e. the coulombic losses are proportional to the useful current-density. This assumption is supported by the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported. In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}. The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption. As a first approximation, a fixed loss term could be placed in the SOC expression in \cref{eqn: Method_SOC_Tracker} instead of the $\sqrt{\eta_C}$ coefficients.

>>>>

\usepackage{graphicx}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
%\setlength{\parindent}{0in}
\usepackage{times}
\usepackage[british]{babel}
\usepackage{algorithm}
\usepackage{algorithmic} \renewcommand{\algorithmicrequire}{\textbf{Given:}} \renewcommand{\algorithmicensure}{\textbf{Returns:}}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\renewcommand \thesection{\arabic{section}}
\usepackage[caption=false]{subfig}
\usepackage[compact]{titlesec}
\usepackage{soul}
\usepackage{textcomp}
\usepackage[section]{placeins}
\usepackage{url}
\usepackage{xspace}
\usepackage{lineno}
\usepackage{hyperref}
    \hypersetup{colorlinks=true,linkcolor=green}
\usepackage{cleveref}

%Added by Diarmid
\usepackage{siunitx}

@@

%% main text
\section{Introduction}
\label{sec:Intro}
The need to reduce carbon dioxide emissions is motivating increasing levels of renewable energy deployment across industrialised nations, and costs for both solar photovoltaic technologies and wind turbines have dropped considerably in recent years. \sol{More citations here}

At a national level, the fluctuating output from these power sources is topped up to match demand primarily by varying output from fossil fuel power thermal plant. In the UK, this is specifically achieved by gas power plant \cite{DraxElectricInsights}. In order to meet the strict CO$_2$ emissions targets for the UK electrical supply by 2050, and bearing in mind that nuclear power capacity is not predicted to grow until after 2030, either carbon capture or significantly greater renewable power penetration will be required of the supply side\cite{FES2018}.

In economic terms, the price of electricity at a given moment is set by the operating costs and  of the most expensive generator in the merit order \cite{Staffell}. As wind and solar have zero marginal cost, increasing penetration of these power sources will reduce electrical price according to a stochastic distribution. At the same time, increasing operating costs for dispatchable fossil generation due to reduced capacity factors and carbon taxation will increase the price in some periods.

Such price variability presents an increasing opportunity for arbitrage by energy storage systems (ESS). At present however, revenue from arbitrage on the electrical price alone is not sufficient to justify BESS investment, even under optimistic economic assumptions \cite{Hu2010}. Hence an important aspect of BESS economics is revenue stacking, where multiple revenue streams are captured by the same system. For example by providing grid reliability services as well as performing market arbitrage. In the UK this is illustrated by the fact that the majority of successful BESS applicants for enhanced frequency response (EFR) contracts also have capacity market contracts \cite{NGDeratingFactors}.

\cite{Sabihuddin2015} classify the services that BESS may provide by the continuous charge/discharge duration, from power quality and regulation ($<$ \SI{1}{\minute}), through bridging power (\SI{1}{\minute} to \SI{1}{\hour}) to energy management ($>$ \SI{1}{hour}). In order to satisfy a given duration a BESS must be specified with the corresponding energy to power ratio.

At the shortest timescale, BESS may receive payments from the grid operator to respond to fluctuations in the AC grid frequency, as in the aforementioned EFR contracts. Such frequency regulation services attract Li-ion systems designed for power, such as the Hornsdale Power Reserve project in South Australia, which is specified for \SI{1}{\hour}\SI{17}{\minute} at its \SI{100}{\mega\watt} rated power output \cite{DOEGlobalESSDatabase}. This energy to power ratio is representative of global Li-ion installation data as collated by \cite{Newbery2018}.

An example of multi-hour energy management is the use of ESS to discharge into peak power demand in a process known as peak shaving. In California, where BESS are being considered as replacement for gas powered peaker plants, a project may receive capacity payments in addition to the sale of energy. However, the regulator de-rates the payment according to a 4 hour target duration, hence a \SI{1}{\mega\watt}/\SI{2}{\mega\watt\hour} BESS would only receive credit for  \SI{0.5}{\mega\watt} of power capacity \cite{NRELPeakingESS}.

A similar de-rating has also been applied recently in the UK capacity market (CM). This was done due to concerns that the average storage duration of the BESS contracted for CM is shorter than the \SI{2}{\hour} that a stress event may last for. \cite{NGDeratingFactors}.

The de-rating developments would disfavour the Li-ion systems that have been deployed so far and favour systems with higher energy to power ratios. Redox flow battery (RFB) systems and promising candidates for multi-hour applications and revenue stacking, as the incremental cost of energy capacity is low, and systems deployed so far have an average energy to power ratio of 3.5 \cite{Newbery2018}.

In order to quantify the economic merits of RFB systems in comparison to other BESS, or indeed ESS, it is necessary to perform techno-economic analyses (TEA) for particular applications. TEA should involve optimisation of the operation of a BESS to maximise revenue (or minimise costs). This should be done at various energy and power ratings (as these are decoupled in a non-hybrid RFB) to find the global optimum, as described by \cite{Oudalov2007}. The optimisation of operation requires determining the schedule, that is, the power output of the battery in each sub-period that maximises benefit without violating energy conservation.

Where a price signal exists optimisation is commonly performed by posing a linear programming (LP) problem \emph{do I dump here a list of all the LP references i will refer to below?}. In this method, the objective function and the constraints on the system are expressed as linear algebraic terms. On the resulting geometric surface, the optimum is guaranteed to lie on a vertex, hence a time-consuming search of the whole surface is not required.

@@

\section{Model formulation}
\label{sec:Model_Formulation}

Pure price arbitrage is chosen as the application for schedule optimisation due to its simplicity, allowing a focus on the treatment of the RFB within the problem. In the following section, an LP formulation for schedule optimisation for maximum arbitrage revenue is first posed in current-density terms to serve as a reference. Then the three augmentations to the model described in \cref{sec:Intro} are introduced. Finally, the RFB specification process is defined.

\section{LP schedule Optimisation in Current-Density Terms}
\label{Model_Formulation_LP_Reference_Current_Density_Terms}

In the literature on RFB schedule optimisation, scheduling problems are posed using either power (\cite{Chen2012,Hu2010,Nguyen2015}) or energy (\cite{Johnston2015,Vaca2017}) as the time-indexed variables. This is not an important difference, as the former only requires that the timestep $\tau$ be included as a coefficient. Here, the power convention is adopted, as it allows the timestep to be varied independently. The objective function for pure arbitrage revenue described by \cite{Hu2010} is adopted here, but adapted so that separate charge and discharge power variables are explicitly defined. Specifying a single power variable with negative values representing charge and positive discharge is not practical for LP optimisation, as the solver will drive the variable to extremes which is likely to be sub-optimal.

@@

\subsection{NLP treatment of voltaic losses}
\label{Model_Formulation_NLP_treatment_voltaic_losses}

A major simplification in the LP problem in terms of power is the assumption of a constant energy efficiency parameter. This assumption also applies to the LP problem in current-density terms introduced in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}, although separate voltaic and coulombic efficiencies parameters are accounted for for. In reality, ohmic (and pseudo-ohmic) over-potential is a linear function of current-density. The faradaic, or activation over-potential, which occurs when the circuit is closed, may be considered constant with respect to current-density \cite{Aaron2011}. \Cref{eqn: Linear_Schedule_Objective_Function_Current_as_Variable} was adapted to include these losses, by subtracting/adding the faradaic over-potential from/to the OCV during discharge/charge, and including the ohmic losses in the form of the classical $P_{Loss} = I^2R$ equation. This results in a non-linear (NLP) objective function for revenue in period $Y$ given by:

@@

where $V_a$ and $ASR$ are scalar parameters representing faradaic over-potential (\si{\volt}) and area specific resistance (\si{\ohm\square\meter}). The placement of $V_a$ and $ASR$ reflects an assumption of symmetry in voltaic losses between charge and discharge, which is discussed further in \cref{analysis_of_assumptions}. The objective function is subject to the same constraints as described in \cref{Model_Formulation_LP_Reference_Current_Density_Terms}.

Although non-linear, \cref{eqn: Non-Linear_Schedule_Objective_Function_Current_as_Variable} the objective function is convex, as it requires maximising a negative quadratic, hence a gradient based algorithmic solver is guaranteed to find the optimum from any starting point.

@@

Both the LP and NLP solutions involve charging the VRFB when the electrical price is low (e.g. 05:00 to 07:00) and discharging it when the price is high (e.g. 10:00 to 13:00). The LP optimal profile involves discharging or charging the system at high current-density in the four hours with the most extreme prices. The NLP optimal schedule also targets the most extreme periods, but spreads the charging/discharging out. This is the consequence of factoring variable ohmic losses in the objective function, where high current-density is penalised by higher losses. Under the LP formulation, the VRFB is predicted to generate \pounds 0.068, whereas the NLP formulation predicts a revenue of \pounds 0.063. In reality however, the voltaic losses that would occur under the LP schedule would reduce the revenue to \pounds 0.051. Across 2017, LP scheduling would have resulted in \pounds 22.61, against a prediction of \pounds 33.16, and NLP scheduling would have yielded \pounds 27.99.

@@

    \item Coulombic efficiency is assumed to be constant with current-density, i.e. the coulombic losses are proportional to the useful current-density. This assumption is supported by the data reported by \cite{Reed2016}, but the behaviour below \SI{1600}{\ampere\per\square\meter} was not reported. In principle shunt current losses are proportional to cell-voltage, as described by \cite{Xing2011}, and hence a function of SOC and over-voltage (via \cref{eqn: Cell_Voltage_Constraint}.\emph{wait, is it the terminal voltage or the OCV that drives the shunt current?} The dependence of coulombic losses on current-density is therefore weaker than implied by the present assumption. As a first approximation, a fixed loss term could be placed in the SOC expression in \cref{eqn: Method_SOC_Tracker} instead of the $\sqrt{\eta_C}$ coefficients.
